---
output: github_document
---

<!-- README.md is generated from README.Rmd. Edit the .Rmd, then run devtools::build_readme(). -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 150
)
library(japanesecolors)
```

# japanesecolors

<!-- badges: start -->
[![R-CMD-check](https://github.com/paulgp/japanesecolors/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/paulgp/japanesecolors/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

`r nrow(palettes())` curated colour palettes in four collections: **Retro**,
**Minimalism**, **Kawaii** and **Avant Garde**.

Every palette is an exported character vector of hex colours, so the API is
just the palette's name.

Available for **R** and **Python**.

## Installation

```r
# install.packages("pak")
pak::pak("paulgp/japanesecolors")
```

```bash
# Python
pip install "japanesecolors[plot]"
```

## Usage

```{r basic}
library(japanesecolors)

kawaii_tri_07
```

Which means palettes work anywhere R takes colours:

```{r barplot, fig.height = 3, fig.width = 7, fig.alt = "Bar chart of four bars coloured with the avantgarde_four_04 palette."}
barplot(c(6, 9, 4, 7), col = avantgarde_four_04, border = NA, axes = FALSE)
```

```{r show, fig.height = 1.3, fig.width = 7, fig.alt = "The seven swatches of minimalism_multi_02, each labelled with its hex value."}
show_palette(minimalism_multi_02)
```

### ggplot2

```{r ggplot, fig.height = 3.5, fig.width = 7, message = FALSE, fig.alt = "Scatter plot of iris sepal measurements, the three species coloured with kawaii_tri_07."}
library(ggplot2)

ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_point(size = 2.5) +
  scale_colour_japanesecolors("kawaii_tri_07") +
  theme_minimal()
```

`scale_fill_japanesecolors()` and `scale_color_japanesecolors()` work the same
way, or pass the vector straight to `scale_colour_manual(values = kawaii_tri_07)`.

## The collections

```{r collections, echo = FALSE}
knitr::kable(collections()[, c("label", "n_palettes", "n_review")],
             col.names = c("Collection", "Palettes", "Provisional"))
```

```{r collection-examples, echo = FALSE, fig.height = 5.2, fig.width = 7, fig.alt = "Two example palettes from each of the four collections, shown as rows of colour swatches."}
op <- par(mfrow = c(8, 1), mar = c(0.2, 0.2, 1.4, 0.2))
for (id in c("retro_four_03", "retro_multi_01",
             "minimalism_four_05", "minimalism_multi_05",
             "kawaii_tri_07", "kawaii_multi_03",
             "avantgarde_four_04", "avantgarde_multi_09")) {
  show_palette(id, labels = FALSE)
}
par(op)
```

## Browsing

IDs are `<collection>_<type>_<nn>`, where type is `bi`, `tri`, `four` or
`multi`. They are stable API.

```{r browse}
palettes("kawaii", type = "tricolor")[, c("id", "n_colors", "status")]

# find one big enough for seven groups
palette_names(n = 7)

# look up by ID when the palette is chosen at runtime
get_palette("minimalism_multi_02")
```

The visual gallery of all `r nrow(palettes())` palettes is on the
[package website](https://paulgp.com/japanesecolors/articles/gallery.html).

## Which ones work in a chart

These are design palettes, chosen to look good together rather than to encode
categories, so most were never meant for data visualisation.
`r sum(palettes()$dataviz_friendly)` of the `r nrow(palettes())` screen well
for it, on two criteria:

- **Colour-vision deficiency** — colours that separate for a trichromat can
  collapse for a dichromat, so each palette is re-measured under simulated
  deuteranopia, protanopia and tritanopia.
- **Perceptual separation** — categories have to read as different at a glance,
  in small marks.

Both use CIEDE2000 on a palette's closest pair, since a palette is only as
readable as the two colours most easily confused. A third check stops a
near-white colour vanishing against the page.

```{r dataviz}
palettes(dataviz_friendly = TRUE, n = 4)[, c("id", "min_delta_e", "min_delta_e_cvd")]
```

The measurements ship with every palette, so you can set your own bar. The
thresholds are calibrated so that the [Okabe-Ito
palette](https://jfly.uni-koeln.de/color/), designed for colour-vision
deficiency, passes — and palettes that separate well in normal vision but
collapse for a dichromat, like ggplot2's default hue scale, do not.

Worth knowing: no palette of five or more colours clears the bar. Keeping
5--8 colours mutually distinguishable under dichromacy is hard, and these
palettes were not built for it.

## Python

The same palettes, generated from the same canonical data, are available for
matplotlib:

```python
import japanesecolors as jc

jc.kawaii_tri_07                  # ['#3D7EB4', '#DF4472', '#E68C5F']
plt.bar(range(3), y, color=jc.kawaii_tri_07)

jc.set_palette("avantgarde_four_04")   # default colour cycle
jc.register_cmaps()                    # then cmap="jc:retro_multi_01"
```

See [`python/README.md`](python/README.md). A test asserts the R and Python
packages carry identical colour values.

## Provenance

Colour values were manually transcribed from the **printed RGB labels** in
*Japanese Color Matching*, edited and published by SendPoints (Sendpoints
Publishing Company Limited), 2022, ISBN 978-988-760-879-0. Hex is computed from
those triplets; nothing was sampled from photographs or converted from CMYK,
and no value was adjusted because another colour looked more plausible.

`japanesecolors` is an independent, unaffiliated project. It is not associated
with, licensed by, endorsed by or approved by SendPoints. It distributes this
project's transcription of RGB/hex colour values and its own metadata — not the
book's text, layouts, photographs, advertisements or other artwork — and is no
substitute for the book itself.

The MIT licence covers this package's code, metadata and documentation. It
confers no rights in the source work.
