---
title: "Palette gallery"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Palette gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
library(japanesecolors)
```

```{css, echo = FALSE}
.rs-controls {
  display: flex; flex-wrap: wrap; gap: .4rem; align-items: center;
  margin: 1rem 0 1.5rem; padding: .75rem; border: 1px solid #e3e3e3;
  border-radius: 8px; background: #fafafa; position: sticky; top: 0; z-index: 10;
}
.rs-controls button {
  font: inherit; font-size: .85rem; padding: .3rem .7rem; cursor: pointer;
  border: 1px solid #ccc; border-radius: 999px; background: #fff; color: #333;
}
.rs-controls button[aria-pressed="true"] { background: #333; color: #fff; border-color: #333; }
.rs-controls label { font-size: .85rem; color: #555; margin-right: .2rem; }
.rs-count { font-size: .85rem; color: #666; margin-left: auto; }

.rs-pal { margin: 0 0 1.1rem; }
.rs-pal-head {
  display: flex; align-items: baseline; gap: .5rem; flex-wrap: wrap;
  font-size: .85rem; margin-bottom: .3rem;
}
.rs-id { font-family: ui-monospace, SFMono-Regular, Menlo, Consolas, monospace; font-weight: 600; }
.rs-name { color: #555; }
.rs-meta { color: #888; margin-left: auto; font-variant-numeric: tabular-nums; }
.rs-flag {
  font-size: .7rem; text-transform: uppercase; letter-spacing: .04em;
  background: #fde68a; color: #713f12; padding: .1rem .4rem; border-radius: 4px;
}
.rs-dv {
  font-size: .7rem; text-transform: uppercase; letter-spacing: .04em;
  background: #cfe8d5; color: #14532d; padding: .1rem .4rem; border-radius: 4px;
}
@media (prefers-color-scheme: dark) {
  .rs-dv { background: #14532d; color: #cfe8d5; }
}
.rs-bar { display: flex; height: 54px; border-radius: 6px; overflow: hidden; border: 1px solid #0000001a; }
.rs-sw { flex: 1 1 0; position: relative; }
.rs-sw span {
  position: absolute; inset: auto 0 4px 0; text-align: center; font-size: .6rem;
  font-family: ui-monospace, SFMono-Regular, Menlo, Consolas, monospace;
  opacity: 0; transition: opacity .12s;
}
.rs-bar:hover .rs-sw span { opacity: .85; }
.rs-empty { color: #888; font-style: italic; }
@media (prefers-color-scheme: dark) {
  .rs-controls { background: #1c1c1c; border-color: #333; }
  .rs-controls button { background: #262626; color: #ddd; border-color: #444; }
  .rs-controls button[aria-pressed="true"] { background: #eee; color: #111; border-color: #eee; }
  .rs-name { color: #bbb; } .rs-meta { color: #999; }
}
```

All `r nrow(palettes())` palettes, grouped by collection. Hover a palette to
read its hex values. Every palette shown here is an exported object: type its
ID into R and you get the vector.

These are design palettes, so most were never meant to encode categories in a
chart. The `r sum(palettes()$dataviz_friendly)` marked **dataviz** stay
separable both in normal vision and under simulated colour-vision deficiency
— filter to them with the *Charts* buttons, or in code with
`palettes(dataviz_friendly = TRUE)`.

```{r gallery, results = "asis"}
esc <- function(x) {
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  gsub(">", "&gt;", x, fixed = TRUE)
}

# black or white label text, whichever reads better on the swatch
ink <- function(hex) {
  v <- grDevices::col2rgb(hex) / 255
  lin <- ifelse(v <= 0.03928, v / 12.92, ((v + 0.055) / 1.055)^2.4)
  lum <- 0.2126 * lin[1, ] + 0.7152 * lin[2, ] + 0.0722 * lin[3, ]
  ifelse(lum > 0.45, "#000", "#fff")
}

palette_html <- function(row) {
  cols <- get_palette(row$id)
  sw <- paste0(sprintf(
    '<div class="rs-sw" style="background:%s"><span style="color:%s">%s</span></div>',
    cols, ink(cols), esc(cols)), collapse = "")

  name <- if (!identical(row$name, row$id)) sprintf('<span class="rs-name">%s</span>', esc(row$name)) else ""
  flag <- if (identical(row$status, "review")) '<span class="rs-flag">provisional</span>' else ""
  dv <- if (isTRUE(row$dataviz_friendly)) {
    sprintf('<span class="rs-dv" title="Separable in normal vision (dE %.0f) and under simulated colour-vision deficiency (dE %.0f)">dataviz</span>',
            row$min_delta_e, row$min_delta_e_cvd)
  } else ""

  # emit as a single line: indented continuation lines would be parsed as a
  # markdown code block rather than as HTML
  sprintf(paste0(
    '<div class="rs-pal" data-collection="%s" data-family="%s" data-status="%s" data-dataviz="%s">',
    '<div class="rs-pal-head"><span class="rs-id">%s</span>%s%s%s',
    '<span class="rs-meta">%s &middot; %d colours</span></div>',
    '<div class="rs-bar">%s</div>',
    '</div>'),
    row$collection, row$family, row$status,
    if (isTRUE(row$dataviz_friendly)) "yes" else "no",
    esc(row$id), name, flag, dv, esc(row$type), row$n_colors, sw)
}

cat('<div class="rs-controls" id="rs-controls">\n')
cat('<label>Collection</label>')
cat('<button data-filter="collection" data-value="all" aria-pressed="true">All</button>')
for (i in seq_len(nrow(collections()))) {
  cl <- collections()[i, ]
  cat(sprintf('<button data-filter="collection" data-value="%s" aria-pressed="false">%s</button>',
              cl$collection, esc(cl$label)))
}
types <- c(bi = "bicolor", tri = "tricolor", four = "four-color", multi = "multicolor")
cat('<label style="margin-left:1rem">Type</label>')
cat('<button data-filter="family" data-value="all" aria-pressed="true">All</button>')
for (code in names(types)) {
  cat(sprintf('<button data-filter="family" data-value="%s" aria-pressed="false">%s</button>',
              code, types[[code]]))
}
cat('<label style="margin-left:1rem">Charts</label>')
cat('<button data-filter="dataviz" data-value="all" aria-pressed="true">All</button>')
cat('<button data-filter="dataviz" data-value="yes" aria-pressed="false">Dataviz-friendly</button>')
cat('<span class="rs-count" id="rs-count"></span>\n')
cat('</div>\n')

cat('<div id="rs-gallery">\n')
for (coll in collections()$collection) {
  rows <- palettes(coll)
  label <- collections()$label[collections()$collection == coll]
  cat(sprintf('<h2 data-collection-heading="%s">%s <span style="font-weight:400;color:#888;font-size:.8em">(%d)</span></h2>\n',
              coll, esc(label), nrow(rows)))
  for (i in seq_len(nrow(rows))) cat(palette_html(rows[i, ]), "\n")
}
cat('</div>\n')
cat('<p class="rs-empty" id="rs-empty" style="display:none">No palettes match that combination.</p>\n')
```

```{js, echo = FALSE}
(function () {
  var state = { collection: "all", family: "all", dataviz: "all" };
  var controls = document.getElementById("rs-controls");
  if (!controls) return;

  function apply() {
    var pals = document.querySelectorAll("#rs-gallery .rs-pal");
    var shown = 0;
    pals.forEach(function (p) {
      var ok = (state.collection === "all" || p.dataset.collection === state.collection) &&
               (state.family === "all" || p.dataset.family === state.family) &&
               (state.dataviz === "all" || p.dataset.dataviz === state.dataviz);
      p.style.display = ok ? "" : "none";
      if (ok) shown++;
    });
    // hide a collection heading when nothing under it survives the filter
    document.querySelectorAll("[data-collection-heading]").forEach(function (h) {
      var any = document.querySelectorAll(
        '#rs-gallery .rs-pal[data-collection="' + h.dataset.collectionHeading + '"]');
      var visible = Array.prototype.filter.call(any, function (p) {
        return p.style.display !== "none";
      }).length;
      h.style.display = visible ? "" : "none";
    });
    document.getElementById("rs-count").textContent =
      shown + (shown === 1 ? " palette" : " palettes");
    document.getElementById("rs-empty").style.display = shown ? "none" : "";
  }

  controls.addEventListener("click", function (e) {
    var btn = e.target.closest("button[data-filter]");
    if (!btn) return;
    state[btn.dataset.filter] = btn.dataset.value;
    controls.querySelectorAll('button[data-filter="' + btn.dataset.filter + '"]')
      .forEach(function (b) { b.setAttribute("aria-pressed", String(b === btn)); });
    apply();
  });

  apply();
})();
```

## Finding a palette from R

```{r find, echo = TRUE}
# all four-colour Avant Garde palettes
palette_names("avantgarde", type = "four-color")

# anything with at least seven colours
palettes(n = 7)[, c("id", "collection", "n_colors")]

# only fully transcribed palettes
nrow(palettes(include_review = FALSE))

# palettes that screen well for charts, widest first
head(palettes(dataviz_friendly = TRUE)[, c("id", "n_colors", "min_delta_e_cvd")])
```
