---
title: "Getting started with japanesecolors"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with japanesecolors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7)
library(japanesecolors)
```

## Palettes are objects

A palette is a character vector of hex colours, exported under its own name:

```{r}
library(japanesecolors)

kawaii_tri_07
```

So it works anywhere R takes colours:

```{r fig.height = 3}
barplot(c(5, 8, 3), col = kawaii_tri_07, border = NA, axes = FALSE)
```

To look at one:

```{r fig.height = 1.3}
show_palette(avantgarde_multi_09)
```

## Palette IDs

IDs have the form `<collection>_<type>_<nn>`:

| Part | Values |
|---|---|
| collection | `retro`, `minimalism`, `kawaii`, `avantgarde` |
| type | `bi` (bicolor), `tri` (tricolor), `four` (four-color), `multi` (multicolor) |
| number | zero-padded, contiguous within each collection and type |

These IDs are stable API. Descriptive names may be added later as aliases; the
numbered IDs will keep working.

## Browsing

`palettes()` returns the catalogue as a data frame:

```{r}
head(palettes()[, c("id", "collection", "type", "n_colors", "status")])
```

It filters by collection, matching type and number of colours:

```{r}
palettes("minimalism", type = "bicolor")[, c("id", "n_colors")]

# anything wide enough for six groups
palette_names(n = 6)
```

`collections()` summarises the four collections:

```{r}
collections()[, c("collection", "label", "n_palettes", "n_review")]
```

## Choosing a palette programmatically

When the palette is decided at runtime, look it up by ID:

```{r}
get_palette("retro_four_03")

# the first n colours, or reversed
get_palette("retro_four_03", n = 2)
get_palette("retro_four_03", reverse = TRUE)
```

Palettes are curated categorical sets and are never stretched. Asking for more
colours than a palette holds is an error, not a recycled or interpolated colour:

```{r error = TRUE}
get_palette("kawaii_tri_07", n = 5)
```

`palettes(n = 5)` will tell you which palettes are big enough.

## ggplot2

```{r fig.height = 3.5, message = FALSE}
library(ggplot2)

ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_point(size = 2.5) +
  scale_colour_japanesecolors("kawaii_tri_07") +
  theme_minimal()
```

```{r fig.height = 3.5}
ggplot(mpg, aes(class, fill = drv)) +
  geom_bar() +
  scale_fill_japanesecolors("avantgarde_tri_06") +
  theme_minimal()
```

All three take `reverse`, `na.value`, and anything else `scale_*_manual()`
accepts:

```{r fig.height = 3.5}
ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_point(size = 2.5) +
  scale_colour_japanesecolors("retro_tri_02", reverse = TRUE, name = "Iris species") +
  theme_minimal()
```

Being plain vectors, they also work with `scale_colour_manual()`, which lets you
pin colours to specific categories:

```{r fig.height = 3.5}
cols <- setNames(kawaii_tri_07, levels(iris$Species))

ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_point(size = 2.5) +
  scale_colour_manual(values = cols) +
  theme_minimal()
```

For packages that expect a palette *function* of `n`, use `japanesecolors_pal()`:

```{r}
pal <- japanesecolors_pal("minimalism_multi_05")
pal(3)
```

## Transcription status

Colour values were read by hand from printed RGB labels. Readings that were not
reliably legible are recorded as provisional:

```{r}
palettes_needing_review()[1:3, c("palette_id", "order", "hex")]
```

They are exported like any other, and `palettes()` includes them by default with
`status == "review"`. For fully transcribed values only:

```{r}
nrow(palettes(include_review = FALSE))
```

See the [package README](https://paulgp.github.io/japanese-colors/) for where
the data came from.
